Aneuploidy is an evolutionary sidetrack on the path to adaptation

### 2022-07-18 ms/ms.tex
short par from AY with intuition

% Abstract
\begin{abstract}
Aneuploidy is common in eukaryotes, often leading to decreased cell growth and fitness. However, evidence from yeast and fungi, as well as human tumour cells, suggests that aneuploidy can be beneficial under stressful conditions and lead to elevated growth rates and adaptation.
Importantly, aneuploidy differs from point mutations in rate, fitness effect, and reversibility.
Here, we develop an evolutionary model for adaptive evolution with both aneuploidy and  mutation, and apply it within an Bayesian inference framework.
We estimate the formation rate and fitness effect of aneuploidy and mutation in an evolutionary experiment in which \yeast adapted to heat stress.
Our results suggest that aneuploidy is more frequent than previously estimated and that experimental observations cannot be explained without aneuploidy.
We do not find evidence for genetic instability in aneuploid cells.
Our results support the hypothesis that aneuploidy is a transient adaptive hypothesis: the experimental populations adapted by first acquiring a chromosome duplication; then by acquiring a beneficial mutation that likely appeared in aneuploid cells, rather than euploid cells; and finally by rapidly losing the extra chromosome to avoid paying the fitness cost of aneuploidy.
\end{abstract}

@@

>>>>

% Abstract
\begin{abstract}
BLA
%Aneuploidy is common in eukaryotes, often leading to decreased cell growth and fitness. However, evidence from yeast and fungi, as well as human tumour cells, suggests that aneuploidy can be beneficial under stressful conditions and lead to elevated growth rates and adaptation.
%Importantly, aneuploidy differs from point mutations in rate, fitness effect, and reversibility.
%Here, we develop an evolutionary model for adaptive evolution with both aneuploidy and  mutation, and apply it within an Bayesian inference framework.
%We estimate the formation rate and fitness effect of aneuploidy and mutation in an evolutionary experiment in which \yeast adapted to heat stress.
%Our results suggest that aneuploidy is more frequent than previously estimated and that experimental observations cannot be explained without aneuploidy.
%We do not find evidence for genetic instability in aneuploid cells.
%Our results support the hypothesis that aneuploidy is a transient adaptive hypothesis: the experimental populations adapted by first acquiring a chromosome duplication; then by acquiring a beneficial mutation that likely appeared in aneuploid cells, rather than euploid cells; and finally by rapidly losing the extra chromosome to avoid paying the fitness cost of aneuploidy.
\end{abstract}

@@

% avihu: At the first stage there are more 2n cells than 2n+1 cells because of the advantage of aneuploidy under heat. Accordingly, at the seconds stage, there are more 2n+1* cells than 2n*. However, in most cases 2n*M dominate the population (and not 2n*A), because the fitness advantage of a refined mutation that emerge in 2n+1 background in hindered by the cost of aneuploidy and by the extra step needed to eliminate the extra copy of the chromosome.
